Mapping the current knowledge in syndemic research applied to Men who have Sex with Men: a Scoping Review

### 2022-04-22 _bookdown_files/04-results.Rmd
Reviews [1st round]

#Load required packages :
library(tidyverse)
library(readr)
library(scales)
library(ggplot2)
library(ggpubr)
library(sunburstR)
library(d3r)
library(english)

#Load the data from the .csv files
source("./scripts/read_data.R", local = knitr::knit_global())

@@

This section summarizes the findings relevant to our research questions. \@ref(tab:RefTabQuant), \@ref(tab:RefTabQual) and \@ref(tab:RefTabRev) in the supplementary material, summarize every reference identified and the variables extracted to answer our research questions for quantitative studies, qualitative studies and reviews, respectively.

## Literature search

The electronic search identified 840 references and our handsearch of reference lists resulted in the addition of the seminal study by Stall and colleague @Stall2003 . After removing duplicates, 305 records were screened for inclusion.

After screening the title and abstract, 124 records were excluded. The two main reasons for exclusion at this stage were that MSM was not the main study population (n=49) and that syndemic was not the main focus of the paper (n=37). 38 studies were excluded because the type of publication did not meet our inclusion criteria.

The full texts of the 181 remaining references were obtained and read; 66 were excluded after this phase. During the previous step, we had decided to tentatively include papers in which the sample was not entirely comprised of MSM. After reviewing the full paper, we decided to exclude studies if the MSM data was aggregated with data from another population (n=11), most commonly transgender women. The rationale for this decision was to keep the focus clearly on MSM.

@@

#Source the scripts used to manipulate the data from the .csv files
source("./scripts/create_year_design.R", local = knitr::knit_global())
source("./scripts/create_location_data.R", local = knitr::knit_global())
source("./scripts/create_sample.R", local = knitr::knit_global())

@@

The publication date of the studies ranges from `r min(year_design$year)` to `r max(year_design$year)` with a marked increase in the number papers published annually from 2013 onward as illustrated in Figure \@ref(fig:PlotYear). There is also an increase in the diversity of the studies, both in the study design used and the continents where the studies were conducted.

@@

However, despite the increasing diversity, the vast majority of the studies were conducted in North America (N = `r .North_America_studies`), with a large predominance of studies in the United States of America (N= `r .USA_studies`). The majority of the remaining studies were then located in Asia (N = `r .Asia_studies`) or in Europe (N = `r .Europe_studies`). South America (N = `r .South_America_studies`) and Africa (N = `r .Africa_studies`) were the least represented continents.

In terms of design, a similar pattern appears, with most of the studies employing a cross-sectional design (N = `r .N_cross_sec`). The number of longitudinal studies has grown steadily since 2014 and represents `r percent(.N_long/nrow(overview_of_studies))` of the papers. The duration of longitudinal studies ranged from `r min(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` to `r max(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` months.

Compared to quantitative studies, qualitative studies are under-represented, with only `r .N_qual` papers. In terms of analysis method, the most common was thematic (content) analysis (N= `r sum(!is.na(summary_qualitative$thematic_content)) + sum(!is.na(summary_qualitative$thematic))`). The rest of the studies consisted of framework analysis (N = `r sum(!is.na(summary_qualitative$framework))`), grounded theory (N = `r sum(!is.na(summary_qualitative$grounded))`), analytic induction (N = `r sum(!is.na(summary_qualitative$analytic_induction))`) and constant comparative analysis (N = `r sum(!is.na(summary_qualitative$comparative))`).

We identified 2 papers employing mixed methods design [@buttram2015; @Halkitis2012]; Buttram et al. used a cross-sectional design for the quantitative part and in-depth interviews analyzed through grounded theory for the qualitative part while Halkitis et al used a cross-sectional quantitative survey and in-depth discovery interviews.

@@

Finally, the median sample size of the studies was `r median(sample$sample_size)` (range : `r min(sample$sample_size)` - `r number(max(sample$sample_size))`) and the mean age of the participants, when reported, ranged from `r min(sample$age, na.rm = T)` to `r round(max(sample$age, na.rm = T), digits = 0)` years.

@@

Fewer than half of the studies (N = `r sum(!is.na(population$subpopulation))` ) focused on a subpopulation of MSM, of those, we identified 7 types of MSM subpopulations : (a) young MSM ; (b) older MSM ; (c) MSM from a racial/ethnic minority ; (d) MSM living with HIV ; (e) MSM engaged in sex work ; (f) Men who have Sex with Men and Women (MSMW) ; and (g) transgender MSM.

\
Among studies focusing on a specific age group, `r sum(population$age == "Young adult")` studies focus on young adults, `r english(sum(population$age == "Older men"))` studies focus on older MSM[@zepf2020; @Halkitis2012] and `r english(sum(population$age == "Adolescent"))` study focuses on adolescents[@Perry].

\
Black MSM (BMSM) are the most studied MSM from a racial/ethnic minority (N= `r sum(population$race_ethnicity == "Black")`) but studies focusing on them still represent only `r percent(sum(population$race_ethnicity == "Black")/104)` of our sample.\
`r english(sum(population$race_ethnicity == "Latino"))` studies focus on Latino MSM (LMSM) and `r english(sum(population$race_ethnicity == "Latino & Black"))` study focus on both BMSM and LMSM[@cassels2020].\

Concerning the serologic status of MSM in syndemic literature, `r sum(population$serologic_status == "HIV-positive")` studies are entirely comprised of a sample of MSM living with HIV. Notably, we didn't identify any studies focusing on racial minority MSM living with HIV.

\
Finally, the least represented subpopulation of MSM in this review are MSM engaged in sex work (N = `r sum(population$sex_work == "Sex workers")`), MSMW (N = `r sum(population$sex_behav == "MSMW")`) and transgender MSM (N = `r sum(population$gender_identity == "Transgender")`).

@@

\
In order to obtain the most comprehensive knowledge of the representation of MSM subpopulations in syndemic literature, we also took into account studies that did not focus on any subpopulation in particular but who presented disaggregated data and analyses for one or more MSM subpopulations.\
`r .N_disaggregated` studies presented such data : `r .N_dis_BMSMW` for Black MSMW[@dyer2020], `r .N_dis_MSMW` for MSMW[@branstrom; @Mustanski2014; @friedman2014; @ferlatte2018], `r .N_dis_HIVMSM` for MSM living with HIV[@Kurtz2012; @ng2020] and `r .N_dis_LBMSM` for Latino and Black MSM[@Mustanski2017].

@@

We identified `r nrow(nodes)` different syndemic conditions in the `r nrow(summary_quantitative)` quantitative studies we included : depression (N = `r nodes$studies[nodes$label == "depression"]`), substance use (N = `r nodes$studies[nodes$label == "substance_use"]`), intimate partner violence [IPV] (N = `r nodes$studies[nodes$label == "IPV"]`), childhood sexual abuse [CSA] (N = `r nodes$studies[nodes$label == "CSA"]`), polysubstance use (N = `r nodes$studies[nodes$label == "polysubstance"]`) , binge drinking (N = `r nodes$studies[nodes$label == "binge"]`), alcohol use disorder [AUD] (N = `r nodes$studies[nodes$label == "AUD"]`), sexual compulsivity (N = `r nodes$studies[nodes$label == "sex_comp"]`), experiences of violence (N = `r nodes$studies[nodes$label == "violence"]`), suicidal thoughts and/or attempts (N = `r nodes$studies[nodes$label == "suicide"]`), sexual risk behaviors (N = `r nodes$studies[nodes$label == "sexual_risk"]`), anxiety (N = `r nodes$studies[nodes$label == "anxiety"]`), substance use disorder [SUD] (N = `r nodes$studies[nodes$label == "substance_use_disorder"]`), alcohol use (N = `r nodes$studies[nodes$label == "alcohol_use"]`), discrimination (N = `r nodes$studies[nodes$label == "discrimination"]`), post-traumatic stress disorder [PTSD] (N = `r nodes$studies[nodes$label == "PTSD"]`), chemsex (N = `r nodes$studies[nodes$label == "chemsex"]`), loneliness (N = `r nodes$studies[nodes$label == "loneliness"]`), incarceration (N = `r nodes$studies[nodes$label == "incarceration"]`), unstable housing (N = `r nodes$studies[nodes$label == "unstable_housing"]`), tobacco use (N = `r nodes$studies[nodes$label == "tobacco_use"]`), general mental distress (N = `r nodes$studies[nodes$label == "distress"]`), childhood abuse (N = `r nodes$studies[nodes$label == "childhood"]`), low social support (N = `r nodes$studies[nodes$label == "social_support"]`), low self-esteem (N = `r nodes$studies[nodes$label == "esteem"]`), internalised homophobia (N = `r nodes$studies[nodes$label == "internalised_homophobia"]`), exchange sex (N = `r nodes$studies[nodes$label == "exchange_sex"]`), sexually transmitted infections [STI] (N = `r nodes$studies[nodes$label == "STI"]`), intravenous drug use [IDU] (N = `r nodes$studies[nodes$label == "IDU"]`), poverty (N = `r nodes$studies[nodes$label == "poverty"]`), unemployment (N = `r nodes$studies[nodes$label == "unemployment"]`), poor healthcare access (N = `r nodes$studies[nodes$label == "healthcare_access"]`), involuntary subordination (N = `r nodes$studies[nodes$label == "subordination"]`), HIV diagnosis (N = `r nodes$studies[nodes$label == "HIV"]`), hostility (N = `r nodes$studies[nodes$label == "hostility"]`), stress (N = `r nodes$studies[nodes$label == "stress"]`), experience of trauma (N = `r nodes$studies[nodes$label == "trauma"]`), alexithymia (N = `r nodes$studies[nodes$label == "alexithymia"]`), attention deficit hyperactivity disorder [ADHD] (N = `r nodes$studies[nodes$label == "ADHD"]`), cognitive escape (N = `r nodes$studies[nodes$label == "escape"]`), frequenting gay social venues (N = `r nodes$studies[nodes$label == "venues"]`), hypersexuality (N = `r nodes$studies[nodes$label == "hypersexuality"]`), impulsivity (N = `r nodes$studies[nodes$label == "impulsivity"]`), poor physical health (N = `r nodes$studies[nodes$label == "physical_health"]`), school bullying (N = `r nodes$studies[nodes$label == "bullying"]`), and sleep disturbance (N = `r nodes$studies[nodes$label == "sleep_disturbance"]`).

We chose to distinguish between alcohol use disorder, binge drinking and alcohol use instead of merging them into a broad "alcohol-related syndemic condition". Some authors also considered "heavy drinking" [@Martinez2016a; @martinez2020] or "heavy alcohol use"[@Mimiaga2015b] but we chose to aggregate these conditions under "binge drinking" for clarity. Similarly, we distinguish between substance use, substance use disorder, intravenous drug use, polysubstance use, marijuana use, tobacco use and chemsex since they differ in terms of potential harm and the context of use.

@@

In order to better understand the connections between the syndemic conditions in the studies we elaborated a network plot (Figure \@ref(fig:SyndemicNetwork)). Nodes represent the syndemic conditions studied in the literature and edges (connectors) are drawn between two nodes when two conditions are studied in the same research paper. Edges between nodes are thicker if the two conditions are frequently studied together and the outline of the nodes represent pie chart of their frequency in syndemic literature. For example, depression, the most studied condition, appears in `r percent((nodes$studies[nodes$label == "depression"])/nrow(summary_quantitative))` of our sample. As such, the node is circled by a chart representing this proportion. Moreover, as a force-directed graph, the location of the nodes gives an indication regarding their importance in terms of frequency and linkage to other conditions. As such, depression occupies a more central position than ADHD which was only studied once.

In this figure, we divided the conditions studied into four categories : mental health (e.g. depression, anxiety, PTSD) in blue, social conditions (e.g. substance use, loneliness) in teal, structural conditions (e.g. unemployment, healthcare access) in orange and physical health (e.g. STI, HIV) in red . The figure shows that structural syndemic conditions and physical health-related syndemic conditions are much less central, therefore much less studied, than social syndemic conditions and mental health-related syndemic conditions which are located towards the center of the figure.

Regarding the edges of the network, we identified `r nrow(edges)` pairs of syndemic conditions. However, nearly half of these pairs of conditions (N = `r nrow(.edges_1)`) appear in only one research paper. On the other hand , the most studied pairs of syndemic conditions were depression and IPV (N = `r edges$weight[edges$from == 1 & edges$to == 4]`), depression and substance use (N = (N = `r edges$weight[edges$from == 1 & edges$to == 2]`), depression and CSA (N = `r edges$weight[edges$from == 1 & edges$to == 10]`), IPV and CSA (N = `r edges$weight[edges$from == 4 & edges$to == 10]`) and depression and polysubstance use (N = `r edges$weight[edges$from == 1 & edges$to == 11]`).

```{r syndemic_network_plot, fig.cap="Network of the syndemic conditions studied in quantitative research. As a force-directed network, the most studied and interconnected conditions are the most central. Moreover, each node is circled by a pie chart of their frequency in the review. The thickness of the edges linking two nodes reflects the number of studies in which the two conditions were studied together.", fig.id="SyndemicNetwork", fig.width=7.45, fig.height=5.32, fig.align='center'}

@@

The next part of our research question was to examine the way the syndemic conditions were studied in the literature. For the most often studied syndemic conditions, we found a high degree of variability in their measurement. Different authors used different scales or criteria and, even when the same scale or criteria was used, the cutoff or reference period differed among studies. For the sake of brevity, we will only provide a detailed description of the measurement method for the five most studied syndemic conditions. Detailed information on the syndemic conditions most impacted by methodology heterogeneity is presented in the supplementary material, part C ( \@ref(tab:DepressionTab) to \@ref(tab:AnxietyTab)).

For depression, `r length(unique(.depression_measurement$depression_scale_criteria))` different scales or criteria were used among the `r nodes$studies[nodes$label == "depression"]` studies in which this condition was studied. The most frequently used scales were the full Center for Epidemiologic Studies-Depression scale [CESD] (N = 26) and the 10-items version of this scale [CESD-10] (N=11), the 9-items version of the Patient Health Questionnaire [PHQ-9] (N=9) and the Beck Depression Inventory [BDI-II] (N = 5).\
Cutoff for these scales varied the most for the CESD (range : 16 to 27) and the PHQ-9 (range : 5 to 15).\
When a scale was not used, the most frequent criteria to assess the presence of depression was the participant being on medication for depression in the past 12 months (N = 4).

For substance use, we looked at the class of substances explicitly screened by the authors and the reference period used. The most frequently screened class of substance was stimulants (i.e. cocaine/crack, amphetamines ; N = `r .substances_prev$sum[["stimulants"]]`), followed by depressants (i.e. GHB/GBL, benzodiazepines without prescriptions ; (N = `r .substances_prev$sum[["depressants"]]`), opioids (i.e. opioids misuse and heroin ; N = `r .substances_prev$sum[["opioids"]]`), marijuana (N = `r .substances_prev$sum[["marijuana"]]`), ecstasy /MDMA (N = `r .substances_prev$sum[["ecstasy"]]`), hallucinogens (i.e. ketamine, psilocybine, phencyclidine ; N = `r .substances_prev$sum[["hallucinogens"]]`), inhalants (nitrous oxyde, Popper ; N = `r .substances_prev$sum[["inhalants"]]`) and new psychoactive substances (i.e. synthetic cannabinoids, cathinones ; N = `r .substances_prev$sum[["NPS"]]`).\
Reference period ranged from the past month to over the lifetime.

For polysubstance use, 18 studies considered that there was polysubstance use when three or more classes of substances were used while nine studies defined polysubstance use as more than one class of substance being used. Furthermore, six studies excluded marijuana from the class of substances screened and two studies excluded stimulants, because stimulants were already screened in another syndemic condition.

For IPV, 5 types of intimate partner violence were identified : physical violence (N = `r .IPV_prevalence$sum[["physical IPV"]]`), psychological violence (N = `r .IPV_prevalence$sum[["psychological IPV"]]`), sexual violence (N = `r .IPV_prevalence$sum[["sexual IPV"]]`), gay-related violence (e.g. threats to reveal the partner's sexual orientation ; N = `r .IPV_prevalence$sum[["gay-related IPV"]]`) and HIV-related violence (e.g. threats to disclose the partner's serologic status ; N = `r .IPV_prevalence$sum[["HIV-related IPV"]]`). Reference periods varied from past month to lifetime.

For CSA, we identified `r length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Definition"]))` different definitions and `r english(length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Scale"])))` scales among the `r nodes$studies[nodes$label == "CSA"]` studies. The various definitions differed on the age at which the sexual intercourse occurred to be qualified as childhood sexual abuse, the age gap between the victim and the perpetrator and whether the sexual intercourse was unwanted. The most frequent definition was Finkelhor's definition[@finkelhor1994] : any sexual experience before 13 years old with someone at least five years older or any sexual experience between 13 and 17 years old with someone at least ten years older (N = 7).

@@

We identified `r ncol(.outcomes) - 1` different outcomes in our sample of quantitative studies. Sexual risk behaviors was the most frequently studied outcome, appearing in `r sum(!is.na(.outcomes$outcome_sexual_risk))` research papers. The second most studied outcome was HIV diagnosis (N = `r sum(!is.na(.outcomes$outcome_HIV))`), followed by STI diagnosis (N = `r sum(!is.na(.outcomes$outcome_STI))`), adherence to antiretroviral therapy (N = `r sum(!is.na(.outcomes$outcome_adherence))`), pre-exposure prophylaxis (PrEP) use (N = `r sum(!is.na(.outcomes$outcome_PrEP))`), healthcare use (N = `r sum(!is.na(.outcomes$outcome_healthcare))`) and the syndemic conditions themselves (N = `r sum(!is.na(.outcomes$outcome_syndemic))`). Less often, researchers also used viral load (N = `r sum(!is.na(.outcomes$outcome_VL))`), suicidality (N = `r sum(!is.na(.outcomes$outcome_suicide))`) or engaging in exchange sex (N = `r sum(!is.na(.outcomes$outcome_exchange_sex))`) as outcomes.

>>>>

#Load required packages:
library(tidyverse)
library(readr)
library(scales)
library(ggplot2)
library(ggpubr)
library(sunburstR)
library(d3r)
library(english)

#Load the data from the.csv files
source("./scripts/read_data.R", local = knitr::knit_global())

@@

This section summarises the findings relevant to our research questions. Table \@ref(tab:RefTabQuant), \@ref(tab:RefTabQual) and \@ref(tab:RefTabRev) in the supplementary material, summarise every reference identified and the variables extracted to answer our research questions for quantitative studies, qualitative studies and reviews, respectively.

## Literature Search

The electronic search identified 840 references and our hand search of reference lists resulted in the addition of the seminal study by Stall and colleague [@Stall2003]. After removing duplicates, 305 records were screened for inclusion.

After screening the title and abstract, 124 records were excluded. The two main reasons for the exclusion at this stage were that MSM were not the main study population (n=49) and that syndemic was not the main focus of the paper (n=37). Thirty-eight studies were excluded because the type of publication did not meet our inclusion criteria.

The full texts of the 181 remaining references were obtained and read; Sixty-six were excluded after this phase. During the previous step, we had decided to tentatively include papers in which the sample was not entirely comprised of MSM. After reviewing the full paper, we decided to exclude studies if the MSM data were aggregated with data from another population (n=11), most commonly transgender women. The rationale for this decision was to keep the focus clearly on MSM.

@@

#Source the scripts used to manipulate the data from the.csv files
source("./scripts/create_year_design.R", local = knitr::knit_global())
source("./scripts/create_location_data.R", local = knitr::knit_global())
source("./scripts/create_sample.R", local = knitr::knit_global())

@@

The publication date of the studies ranges from `r min(year_design$year)` to `r max(year_design$year)` with a marked increase in the number of papers published annually from 2013 onward as illustrated in Figure \@ref(fig:PlotYear). There is also an increase in the diversity of the studies, both in the study design used and the continents where the studies were conducted.

@@

However, despite the increasing diversity, the vast majority of the studies were conducted in North America (N = `r .North_America_studies`), with a large predominance of studies in the USA (N= `r .USA_studies`). The majority of the remaining studies were then located in Asia (N = `r .Asia_studies`) or in Europe (N = `r .Europe_studies`). South America (N = `r .South_America_studies`) and Africa (N = `r .Africa_studies`) are the least represented continents.

In terms of design, a similar pattern appears, with most of the studies employed a cross-sectional design (N = `r .N_cross_sec`). The number of longitudinal studies has grown steadily since 2014 and represents `r percent(.N_long/nrow(overview_of_studies))` of the papers. The duration of longitudinal studies ranged from `r min(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` to `r max(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` months.

Compared to quantitative studies, qualitative studies are under-represented, with only `r .N_qual` papers. In terms of analysis method, the most common is thematic (content) analysis (N= `r sum(!is.na(summary_qualitative$thematic_content)) + sum(!is.na(summary_qualitative$thematic))`). The rest of the studies consists of framework analysis (N = `r sum(!is.na(summary_qualitative$framework))`), grounded theory (N = `r sum(!is.na(summary_qualitative$grounded))`), analytic induction (N = `r sum(!is.na(summary_qualitative$analytic_induction))`) and constant comparative analysis (N = `r sum(!is.na(summary_qualitative$comparative))`).

We identified 2 papers employing mixed methods design [@buttram2015; @Halkitis2012]; Buttram et al. used a cross-sectional design for the quantitative part and in-depth interviews analysed through grounded theory for the qualitative part while Halkitis et al. used a cross-sectional quantitative survey and in-depth discovery interviews.

@@

Finally, the median sample size of the studies is `r median(sample$sample_size)` (range: `r min(sample$sample_size)` - `r number(max(sample$sample_size))`) and the mean age of the participants, when reported, ranges from `r min(sample$age, na.rm = T)` to `r round(max(sample$age, na.rm = T), digits = 0)` years.

@@

Less than half of the studies (N = `r sum(!is.na(population$subpopulation))` ) focus on a subpopulation of MSM; of those, we identified seven types of MSM subpopulations: (a) young MSM; (b) older MSM; (c) MSM from a racial/ethnic minority; (d) MSM living with HIV; (e) MSM engaged in sex work; (f) Men who have Sex with Men and Women (MSMW); and (g) transgender MSM.\
Among studies focusing on a specific age group, `r sum(population$age == "Young adult")` studies focus on young adults, `r english(sum(population$age == "Older men"))` studies focus on older MSM [@zepf2020; @Halkitis2012] and `r english(sum(population$age == "Adolescent"))` study focuses on adolescents [@Perry].\
Black MSM (BMSM) are the most studied MSM from a racial/ethnic minority (N= `r sum(population$race_ethnicity == "Black")`) but studies focusing on them still represent only `r percent(sum(population$race_ethnicity == "Black")/104)` of our sample.\
`r english(sum(population$race_ethnicity == "Latino"))` studies focus on Latino MSM (LMSM) and `r english(sum(population$race_ethnicity == "Latino & Black"))` study focus on both BMSM and LMSM [@cassels2020].

Concerning the serologic status of MSM in syndemic literature, `r sum(population$serologic_status == "HIV-positive")` studies are entirely comprised of a sample of MSM living with HIV. Notably, we didn't identify any studies focusing on racial minority MSM living with HIV.\
Finally, the least represented subpopulation of MSM in this review are MSM engaged in sex work (N = `r sum(population$sex_work == "Sex workers")`), MSMW (N = `r sum(population$sex_behav == "MSMW")`) and transgender MSM (N = `r sum(population$gender_identity == "Transgender")`).

@@

\
In order to obtain the most comprehensive knowledge of the representation of MSM subpopulations in syndemic literature, we also took into account studies that do not focus on any subpopulation in particular but present disaggregated data and analyses for one or more MSM subpopulations.\
`r .N_disaggregated` studies present such data: `r .N_dis_BMSMW` for Black MSMW [@dyer2020], `r .N_dis_MSMW` for MSMW [@branstrom; @Mustanski2014; @friedman2014; @ferlatte2018], `r .N_dis_HIVMSM` for MSM living with HIV [@Kurtz2012; @ng2020] and `r .N_dis_LBMSM` for Latino and Black MSM [@Mustanski2017].

@@

We identified `r nrow(nodes)` different syndemic conditions in the `r nrow(summary_quantitative)` quantitative studies we included.

The health-related conditions we identified are : depression (N = `r nodes$studies[nodes$label == "depression"]`), alcohol use disorder (AUD) (N = `r nodes$studies[nodes$label == "AUD"]`), sexual compulsivity (N = `r nodes$studies[nodes$label == "sex_comp"]`), suicidal thoughts and/or attempts (N = `r nodes$studies[nodes$label == "suicide"]`), anxiety (N = `r nodes$studies[nodes$label == "anxiety"]`), substance use disorder (SUD) (N = `r nodes$studies[nodes$label == "substance_use_disorder"]`), post-traumatic stress disorder (PTSD) (N = `r nodes$studies[nodes$label == "PTSD"]`), general mental distress (N = `r nodes$studies[nodes$label == "distress"]`), internalised homophobia (N = `r nodes$studies[nodes$label == "internalised_homophobia"]`), low self-esteem (N = `r nodes$studies[nodes$label == "esteem"]`), sexually transmitted infections (STI) (N = `r nodes$studies[nodes$label == "STI"]`), involuntary subordination (N = `r nodes$studies[nodes$label == "subordination"]`), HIV diagnosis (N = `r nodes$studies[nodes$label == "HIV"]`), hostility (N = `r nodes$studies[nodes$label == "hostility"]`), stress (N = `r nodes$studies[nodes$label == "stress"]`), experience of trauma (N = `r nodes$studies[nodes$label == "trauma"]`), alexithymia (N = `r nodes$studies[nodes$label == "alexithymia"]`), attention deficit hyperactivity disorder (ADHD) (N = `r nodes$studies[nodes$label == "ADHD"]`), cognitive escape (N = `r nodes$studies[nodes$label == "escape"]`), hypersexuality (N = `r nodes$studies[nodes$label == "hypersexuality"]`), impulsivity (N = `r nodes$studies[nodes$label == "impulsivity"]`), poor physical health (N = `r nodes$studies[nodes$label == "physical_health"]`) and sleep disturbance (N = `r nodes$studies[nodes$label == "sleep_disturbance"]`).

Alongside these health-related conditions, we also identified numerous social conditions: substance use (N = `r nodes$studies[nodes$label == "substance_use"]`), intimate partner violence (IPV) (N = `r nodes$studies[nodes$label == "IPV"]`), childhood sexual abuse (CSA) (N = `r nodes$studies[nodes$label == "CSA"]`), polysubstance use (N = `r nodes$studies[nodes$label == "polysubstance"]`), binge drinking (N = `r nodes$studies[nodes$label == "binge"]`), experiences of violence (N = `r nodes$studies[nodes$label == "violence"]`), sexual risk behaviours (N = `r nodes$studies[nodes$label == "sexual_risk"]`), alcohol use (N = `r nodes$studies[nodes$label == "alcohol_use"]`), discrimination (N = `r nodes$studies[nodes$label == "discrimination"]`), chemsex (N = `r nodes$studies[nodes$label == "chemsex"]`), loneliness (N = `r nodes$studies[nodes$label == "loneliness"]`), incarceration (N = `r nodes$studies[nodes$label == "incarceration"]`), unstable housing (N = `r nodes$studies[nodes$label == "unstable_housing"]`), tobacco use (N = `r nodes$studies[nodes$label == "tobacco_use"]`), childhood abuse (N = `r nodes$studies[nodes$label == "childhood"]`), low social support (N = `r nodes$studies[nodes$label == "social_support"]`), exchange sex (N = `r nodes$studies[nodes$label == "exchange_sex"]`), intravenous drug use (IDU) (N = `r nodes$studies[nodes$label == "IDU"]`), poverty (N = `r nodes$studies[nodes$label == "poverty"]`), unemployment (N = `r nodes$studies[nodes$label == "unemployment"]`), poor healthcare access (N = `r nodes$studies[nodes$label == "healthcare_access"]`), frequenting gay social venues (N = `r nodes$studies[nodes$label == "venues"]`) and school bullying (N = `r nodes$studies[nodes$label == "bullying"]`).

We chose to distinguish between AUD, binge drinking and alcohol use instead of merging them into a broad 'alcohol-related syndemic condition'. Some authors also considered 'heavy drinking' [@Martinez2016a; @martinez2020] or 'heavy alcohol use' [@Mimiaga2015b] but we chose to aggregate these conditions under 'binge drinking' for clarity. Similarly, we distinguish between substance use, substance use disorder, intravenous drug use, polysubstance use, marijuana use, tobacco use and chemsex since they differ in terms of potential harm and the context of use.

@@

In order to better understand the connections between the syndemic conditions in the studies, we elaborated a network plot (Figure \@ref(fig:SyndemicNetwork)). Nodes represent the syndemic conditions studied in the literature and edges (connectors) are drawn between two nodes when two conditions are studied in the same research paper. Edges between nodes are thicker if the two conditions are frequently studied together and the outline of the nodes represent pie chart of their frequency in syndemic literature. For example, depression, the most studied condition, appears in `r percent((nodes$studies[nodes$label == "depression"])/nrow(summary_quantitative))` of our sample. As such, the node is circled by a chart representing this proportion. Moreover, as a force-directed graph, the location of the nodes gives an indication regarding their importance in terms of frequency and linkage to other conditions. As such, depression occupies a more central position than ADHD, which was only studied once.

In this figure, we divided the conditions studied into four categories: mental health (e.g. depression, anxiety, PTSD) in blue, social conditions (e.g. substance use, loneliness) in teal, structural conditions (e.g. unemployment, healthcare access) in orange and physical health (e.g. STI, HIV) in red. The figure shows that structural syndemic conditions and physical health-related syndemic conditions are much less central, therefore much less studied, than social syndemic conditions and mental health-related syndemic conditions which are located toward the center of the figure.

Regarding the edges of the network, we identified `r nrow(edges)` pairs of syndemic conditions. However, nearly half of these pairs of conditions (N = `r nrow(.edges_1)`) appear in only one research paper. On the other hand, the most studied pairs of syndemic conditions were depression and IPV (N = `r edges$weight[edges$from == 1 & edges$to == 4]`), depression and substance use (N = (N = `r edges$weight[edges$from == 1 & edges$to == 2]`), depression and CSA (N = `r edges$weight[edges$from == 1 & edges$to == 10]`), IPV and CSA (N = `r edges$weight[edges$from == 4 & edges$to == 10]`) and depression and polysubstance use (N = `r edges$weight[edges$from == 1 & edges$to == 11]`).

```{r syndemic_network_plot, fig.cap="Network of the syndemic conditions studied in quantitative research. As a force-directed network, the most studied and interconnected conditions are the most central. Additionally, each node is circled by a pie chart of their frequency in the review. The thickness of the edges linking two nodes reflects the number of studies in which the two conditions were studied together.", fig.id="SyndemicNetwork", fig.width=7.45, fig.height=5.32, fig.align='center'}

@@

The next part of our research question was to examine the way the syndemic conditions were studied in the literature. For the most often studied syndemic conditions, we found a high degree of variability in their measurement. Different authors used different scales or criteria and, even when the same scale or criteria were used, the cutoff or reference period differed among studies. For the sake of brevity, we will only provide a detailed description of the measurement method for the five most studied syndemic conditions. Detailed information on the syndemic conditions most impacted by methodology heterogeneity is presented in the supplementary material, part C ( \@ref(tab:DepressionTab) to \@ref(tab:AnxietyTab)).

For depression, `r length(unique(.depression_measurement$depression_scale_criteria))` different scales or criteria are used among the `r nodes$studies[nodes$label == "depression"]` studies in which this condition was studied. The most frequently used scales are the full Center for Epidemiologic Studies-Depression scale (CESD) (N = 26) and the 10-item version of this scale (CESD-10) (N=11), the 9-item version of the Patient Health Questionnaire (PHQ-9) (N=9) and the Beck Depression Inventory (BDI-II) (N = 5).\
Cutoff for these scales varied the most for the CESD (range: 16 to 27) and the PHQ-9 (range: 5 to 15).\
When a scale is not used, the most frequent criteria to assess the presence of depression is the participant being on medication for depression in the past 12 months (N = 4).

For substance use, we looked at the class of substances explicitly screened by the authors and the reference period used. The most frequently screened class of substance is stimulants (i.e. cocaine/crack, amphetamines; N = `r .substances_prev$sum[["stimulants"]]`), followed by depressants (i.e. GHB/GBL, benzodiazepines without prescriptions; (N = `r .substances_prev$sum[["depressants"]]`), opioids (i.e. opioids misuse and heroin; N = `r .substances_prev$sum[["opioids"]]`), marijuana (N = `r .substances_prev$sum[["marijuana"]]`), ecstasy /MDMA (N = `r .substances_prev$sum[["ecstasy"]]`), hallucinogens (i.e. ketamine, psilocybin, phencyclidine; N = `r .substances_prev$sum[["hallucinogens"]]`), inhalants (nitrous oxide, popper; N = `r .substances_prev$sum[["inhalants"]]`) and new psychoactive substances (i.e. synthetic cannabinoids, cathinones; N = `r .substances_prev$sum[["NPS"]]`).\
Reference periods range from the past month to over the lifetime.

For polysubstance use, 18 studies consider that there is polysubstance use when three or more classes of substances are used while nine studies define polysubstance use as more than one class of substance being used. Furthermore, six studies exclude marijuana from the class of substances screened and two studies exclude stimulants, because stimulants were already screened in another syndemic condition.

Five types of IPV were identified: physical violence (N = `r .IPV_prevalence$sum[["physical IPV"]]`), psychological violence (N = `r .IPV_prevalence$sum[["psychological IPV"]]`), sexual violence (N = `r .IPV_prevalence$sum[["sexual IPV"]]`), gay-related violence (e.g. threats to reveal the partner's sexual orientation; N = `r .IPV_prevalence$sum[["gay-related IPV"]]`) and HIV-related violence (e.g. threats to disclose the partner's serologic status; N = `r .IPV_prevalence$sum[["HIV-related IPV"]]`). Reference periods vary from the past month to over the lifetime.

For CSA, we identified `r length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Definition"]))` different definitions and `r english(length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Scale"])))` scales among the `r nodes$studies[nodes$label == "CSA"]` studies. The various definitions differ regarding the age at which sexual intercourse occurred to be qualified as childhood sexual abuse, the age gap between the victim and the perpetrator and whether the sexual intercourse was unwanted. The most frequent definition is Finkelhor's definition [@finkelhor1994]: any sexual experience before 13 years old with someone at least five years older or any sexual experience between 13 and 17 years old with someone at least ten years older (N = 7).

@@

We identified `r ncol(.outcomes) - 1` different outcomes in our sample of quantitative studies. Sexual risk behaviour is the most frequently studied outcome, appearing in `r sum(!is.na(.outcomes$outcome_sexual_risk))` research papers. The second most studied outcome is HIV diagnosis (N = `r sum(!is.na(.outcomes$outcome_HIV))`), followed by STI diagnosis (N = `r sum(!is.na(.outcomes$outcome_STI))`), adherence to antiretroviral therapy (N = `r sum(!is.na(.outcomes$outcome_adherence))`), pre-exposure prophylaxis (PrEP) use (N = `r sum(!is.na(.outcomes$outcome_PrEP))`), healthcare use (N = `r sum(!is.na(.outcomes$outcome_healthcare))`) and the syndemic conditions themselves (N = `r sum(!is.na(.outcomes$outcome_syndemic))`). Less often, researchers also use viral load (N = `r sum(!is.na(.outcomes$outcome_VL))`), suicidality (N = `r sum(!is.na(.outcomes$outcome_suicide))`) or engaging in exchange sex (N = `r sum(!is.na(.outcomes$outcome_exchange_sex))`) as outcomes.
